---
title: "Final Assignment"
output: html_document
date: "2024-04-23"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


#final Project 
#Rameshwaree Sharma 
#AIRBNB DATASET

```{r}
colours <- c("blue", "red", "green", "yellow", "orange")
```

```{r}
install.packages("splitstackshape")
```

```{r}
library(splitstackshape)
```


```{r}
airbnb <- read.csv("/Users/rameshwaree/Desktop/AirB_NYC_2019\ .csv")
```

```{r}
head(airbnb)
```


```{r}
str(airbnb)
```

```{r}
any(is.na(airbnb))
```

```{r}
sum(is.na(airbnb))
```
#There were 10052 Null values which we removed 
```{r}
airbnb<- na.omit(airbnb)
data <- unique(airbnb)
remaining_observations <- nrow(airbnb)
print(remaining_observations)
```


```{r}
unique(airbnb$price)
```

```{r}
unique(airbnb$availability_365)
```

```{r}
unique(airbnb$neighbourhood_group)
```

```{r}
paste("Price Minimum:", min(airbnb$price), "| Price Maximum:", max(airbnb$price))
```

```{r}
mean_price <- tapply(airbnb$price, airbnb$availability_365, FUN = mean)
median_price <- tapply(airbnb$price, airbnb$availability_365, FUN = median)
sd_price <- tapply(airbnb$price, airbnb$availability_365, FUN = sd)

# Combining results into a data frame
summary_stats <- data.frame(price = unique(airbnb$availability_365),
                            mean_price = mean_price,
                            median_price = median_price,
                            sd_price = sd_price)
print(summary_stats)
```

```{r}
airbnb$log_price <- log(airbnb$price)
```

```{r}
summary(airbnb)
```
#EDA
```{r}
# Splitting the data by 'neighbourhood_group'
neighbourhoods <- unique(airbnb$neighbourhood_group)

# Create separate histograms for each neighbourhood group
par(mfrow = c(2, 3))  # Setting up a 2x3 layout for multiple plots
for (neighbourhood in neighbourhoods) {
  hist(airbnb$log_price[airbnb$neighbourhood_group == neighbourhood],
       main = paste("Histogram of Listing Prices in", neighbourhood),
       xlab = "Listing Price")
}

```

```{r}
boxplot(log_price ~ neighbourhood_group, data = airbnb,
        main = "Boxplot of Listing Prices by Neighbourhood Group",
        xlab = "Neighbourhood Group", ylab = "Listing Price")
```
```{r}
boxplot(airbnb$minimum_nights, main = "Boxplot of minimum_nights")
```
```{r}
# Q1 <- quantile(airbnb$minimum_nights, 0.25)
#Q3 <- quantile(airbnb$minimum_nights, 0.75)
#IQR <- Q3 - Q1
# lower_bound <- Q1 - 1.5 * IQR
#upper_bound <- Q3 + 1.5 * IQR
# cleaned_data <- airbnb[airbnb$minimum_nights >= lower_bound & airbnb$minimum_nights <= upper_bound, ]
```


```{r}
airbnb <- airbnb[airbnb$log_price > 0, ]
summary(airbnb$log_price)
hist(airbnb$log_price, main = "Distribution of Prices", xlab = "Price")
```
Positive skewed 




```{r}
hist(subset(airbnb$log_price, airbnb$log_price <= 1000),
     col = "Pink",       
     breaks = 70,        
     border = "black",   
     main = "Price <= 1000 | Histogram",  
     xlab = "Price",     
     ylab = "Frequency", 
     ylim = c(0, 1000))  

```

```{r}
plot(airbnb$log_price ~ airbnb$number_of_reviews  , main = "Price vs.number_of_reviews  ", xlab = "number_of_reviews ", ylab = "Price")

```

```{r}
plot(airbnb$log_price ~ airbnb$availability_365  , main = "Price vs. availability_365  ", xlab = "availability_365 ", ylab = "Price")

```

```{r}
plot(airbnb$price <= 1000 ~ airbnb$availability_365  , main = "Price vs. availability_365  ", xlab = "availability_365 ", ylab = "Price")

```

```{r}
Neighbourhood <- xtabs(~ neighbourhood_group, data = airbnb)
orderIDX <- order(Neighbourhood, decreasing = TRUE)
yLimit <- range(pretty(range(Neighbourhood)))
XX <- barplot(Neighbourhood[orderIDX], ylim = yLimit, axes = FALSE, names.arg = FALSE, col = colours[1], border = colours[1])
axis(2, at = pretty(yLimit), tck = -.015, lwd =2, labels = FALSE)
mtext(pretty(yLimit),2,at =pretty(yLimit), line = .4, las = 1)
mtext("Frequency", 2, font = 2, line = 3)
mtext(rownames(Neighbourhood)[orderIDX], 1, at = XX, line = .1, padj = c(1,0), cex = 1)

```
##From the above graph, we can tell that Manhattan and Brooklyn have the highest number of the listing 

#Research Question 

##How does the inclusion of neighborhood and property type impact the accuracy of predictive models compared to  number of reviews, and availability of amenities, 
###NULL 
###ALTERNATIVE

```{r}
cor(airbnb$price <= 1000, airbnb$number_of_reviews )
```

```{r}
cor(airbnb$price <= 1000, airbnb$availability_365 )

```

```{r}
price_percentiles <- quantile(airbnb$price, probs = seq(0, 1, by = 0.20), na.rm = TRUE)
print(price_percentiles)
```

```{r}
avail_percentiles <- quantile(airbnb$availability_365, probs = seq(0, 1, by = 0.20), na.rm = TRUE)
print(avail_percentiles)
```

```{r}
min(airbnb$price)
```

```{r}
max(airbnb$price)
```

```{r}
mean(airbnb$price)
```

```{r}
mean(airbnb$availability_365)
```

#REGRESSION 

```{r}
out <- boxplot(airbnb$price, plot = FALSE)
airbnb3 <- airbnb[!(airbnb$price %in% out$out), ]
```


```{r}
set.seed(42)
trainIDX <- sample(c(TRUE, FALSE), size = dim(airbnb)[1], replace = TRUE, prob = c(.80,.20))
airbnb.Train <- airbnb[trainIDX,]
airbnb.Test <- airbnb[!trainIDX,]
```

```{r}
theTwoSamples <- splitstackshape::stratified(airbnb, group = "neighbourhood_group", size = .6, replace = FALSE, bothSets = TRUE)
airbnb.Train <- as.data.frame(theTwoSamples$SAMP1)
testValidate <- splitstackshape::stratified(theTwoSamples$SAMP2, group = "neighbourhood_group", size = .5, replace = FALSE, bothSets = TRUE)
airbnb.Test <- as.data.frame(testValidate$SAMP1)
airbnb.Validate <- as.data.frame(testValidate$SAMP2)
```

```{r}
pTrain <- nrow(airbnb.Train)/nrow(airbnb)
pTest <- nrow(airbnb.Test)/nrow(airbnb)
pValidate <- nrow(airbnb.Validate)/nrow(airbnb)

data.frame(Train = pTrain, Test = pTest, Validate = pValidate)
```

```{r}
#encoded_data <- cbind(airbnb, model.matrix(~ neighbourhood_group + room_type - 1, data = airbnb))
```

```{r}
theModel <- lm(log_price ~ minimum_nights + longitude +latitude + number_of_reviews + availability_365, data = airbnb.Train)
summary(theModel)
```


```{r}
theModel1 <- lm(log_price ~ neighbourhood_group + room_type + number_of_reviews + availability_365, data = airbnb.Train)
summary(theModel1)
```

```{r}
#plot(theModel)
```


```{r}

theVariables <- c( "minimum_nights", "longitude","latitude", "number_of_reviews","availability_365")

par(mfcol = c(2,3), mai = c(.55,.25,.1,.1), omi = c(.3,.3,.3,.3))
yLim <- range(pretty(range(airbnb.Train$log_price)))

for (i in 1:length(theVariables)){
  xLim <- range(pretty(range(airbnb.Train[, theVariables[i]])))
  plot(x = NULL, y = NULL, xlim = xLim, ylim = yLim, axes = FALSE, ann = FALSE)
  axis(1, at = pretty(xLim),lwd = 2, tck = -.015, labels = FALSE)
  axis(2, at = pretty(yLim),lwd = 2, tck = -.015, labels = FALSE)
  mtext(pretty(xLim), 1, at = pretty(xLim), line = .5)
  mtext(theVariables[i],1, font =2, line = 1.75)
  if (i == 1 | i == 2){
    mtext(pretty(yLim), 2, at = pretty(yLim), line = .4, las = 1)
    mtext("price_log", 2, font = 2, line = 2)
  }
  points(airbnb.Train[, theVariables[i]], airbnb.Train$log_price, pch = 21, bg = colours[2] , col = colours[1])
  abline(lm(as.formula(paste0("log_price ~ ", theVariables[i]))
, data = airbnb.Train), lwd = 3, col = "gray75")
}
```
#Normality of the residuals 

```{r}
theResiduals <- theModel$residuals
```

```{r}
axisValues <- qqnorm(theResiduals, plot.it = FALSE)
xLim <- range(pretty(range(axisValues$x)))
yLim <-  range(pretty(range(axisValues$y)))

qqnorm(theResiduals, xlim = xLim, ylim = yLim, pch = 21, col = colours[2], bg = colours[1], cex = 1.5, axes = FALSE, ann = FALSE)
axis(1, at = pretty(xLim), lwd = 2, tck = -.015, labels = FALSE, pos = 0)
axis(1, at = pretty(xLim), lwd = 2, tck = .015, labels = FALSE, pos = 0)
axis(2, at = pretty(yLim), lwd = 2, tck = -.015, labels = FALSE, pos = 0)
axis(2, at = pretty(yLim), lwd = 2, tck = .015, labels = FALSE, pos = 0)

# Text Labelling in the plot
text(pretty(xLim)[!pretty(xLim) == 0], -.5, pretty(xLim)[!pretty(xLim) == 0])
text(-.5, pretty(yLim)[!pretty(yLim) == 0], pretty(yLim)[!pretty(yLim) == 0])
text(max(pretty(xLim)), -1, "Theoretical Quantiles", font = 2, adj = 1)
text(-5, min(pretty(yLim)), "Sample Quantiles", font = 2, adj = 0, srt = 90)

qqline(theResiduals, lwd=3, col ="gray75")
```
#Homoscedasticity
```{r}
xLim <- range(pretty(range(theModel$fitted.values)))
yLim <- range(pretty(range(theResiduals)))
plot(x = NULL, y = NULL, xlim = xLim, ylim = yLim, axes = FALSE, ann = FALSE)
axis(1, at = pretty(xLim), lwd = 2, tck = -.015, labels = FALSE, pos=0)
axis(1, at = pretty(xLim), lwd = 2, tck = .015, labels = FALSE, pos=0)
axis(2, at = pretty(yLim), lwd = 2, tck = .015, labels = FALSE)
axis(2, at = pretty(yLim), lwd = 2, tck = -.015, labels = FALSE)
mtext(pretty(yLim),2,at = pretty(yLim), line = .4, las = 1)
mtext("Residuals", 2, font = 2, line= 1.5)

text(pretty(xLim), -.5, pretty(xLim))
text(max(pretty(xLim)), -1, "Fitted Values", font =2, adj=1)
points(theModel$fitted.values, theResiduals, pch = 21, col = colours[2], bg = colours[1], cex = 1.5)
```
#Multicollinearity
#influentialvalue and outliers
```{r}
standardRez <- rstandard(theModel)
yLim <- range(pretty(range(standardRez)))
plot(x = NULL, y = NULL, xlim = c(1,length(standardRez)), ylim = yLim, axes = FALSE, ann = FALSE)
axis(2, at = pretty(yLim), lwd = 2, tck = -.015, labels = FALSE)
mtext(pretty(yLim), 2, line = .25, at = pretty(yLim), las = 1)
mtext("Standardized Residuals", 2, font = 2, line = 1.5)

points(1:length(standardRez), standardRez, pch = 21, col = colours[2], bg = colours[1], cex = 1.5)
abline(h = -3, lwd = 3, col = "gray75")
abline(h = 3, lwd = 3, col = "gray75")
```
```{r}
outliersIDX <- abs(standardRez) >3
```

```{r}
cookD <- cooks.distance(theModel)

yLim <- range(pretty(range(cookD)))
plot(x = NULL, y = NULL, xlim = c(1,length(cookD)), ylim = yLim, axes = FALSE, ann = FALSE)
axis(2, at = pretty(yLim), lwd = 2, tck = -.015, labels = FALSE)
mtext(pretty(yLim), 2, line = .25, at = pretty(yLim), las = 1)
mtext("Cook's Distance", 2, font = 2, line = 1.5)

points(1:length(cookD), cookD, pch = 21, col = colours[1], bg = colours[1], cex = 1.5)
```
```{r}
cutOff <- mean(cookD)+3*sd(cookD)
cookIDX <- cookD > cutOff
```

```{r}
diagnosticIDX <- as.logical(outliersIDX + cookIDX)
airbnb.Train2 <- airbnb.Train[!diagnosticIDX,]
```

```{r}
theModel.Clean <- lm(log_price ~  minimum_nights + longitude +latitude + number_of_reviews + availability_365, data = airbnb.Train2)
```

```{r}

theVariables <- c( "minimum_nights", "longitude","latitude", "number_of_reviews","availability_365")

par(mfcol = c(2,3), mai = c(.55,.25,.1,.1), omi = c(.3,.3,.3,.3))
yLim <- range(pretty(range(airbnb.Train2$log_price)))

for (i in 1:length(theVariables)){
  xLim <- range(pretty(range(airbnb.Train2[, theVariables[i]])))
  plot(x = NULL, y = NULL, xlim = xLim, ylim = yLim, axes = FALSE, ann = FALSE)
  axis(1, at = pretty(xLim),lwd = 2, tck = -.015, labels = FALSE)
  axis(2, at = pretty(yLim),lwd = 2, tck = -.015, labels = FALSE)
  mtext(pretty(xLim), 1, at = pretty(xLim), line = .5)
  mtext(theVariables[i],1, font =2, line = 1.75)
  if (i == 1 | i == 2){
    mtext(pretty(yLim), 2, at = pretty(yLim), line = .4, las = 1)
    mtext("price_log", 2, font = 2, line = 2)
  }
  points(airbnb.Train2[, theVariables[i]], airbnb.Train2$log_price, pch = 21, bg = colours[2] , col = colours[1])
  abline(lm(as.formula(paste0("log_price ~ ", theVariables[i]))
, data = airbnb.Train2), lwd = 3, col = "gray75")
}
```

```{r}
axisValues <- qqnorm(theResiduals, plot.it = FALSE)
xLim <- range(pretty(range(axisValues$x)))
yLim <-  range(pretty(range(axisValues$y)))

qqnorm(theResiduals, xlim = xLim, ylim = yLim, pch = 21, col = colours[2], bg = colours[1], cex = 1.5, axes = FALSE, ann = FALSE)
axis(1, at = pretty(xLim), lwd = 2, tck = -.015, labels = FALSE, pos = 0)
axis(1, at = pretty(xLim), lwd = 2, tck = .015, labels = FALSE, pos = 0)
axis(2, at = pretty(yLim), lwd = 2, tck = -.015, labels = FALSE, pos = 0)
axis(2, at = pretty(yLim), lwd = 2, tck = .015, labels = FALSE, pos = 0)

# Text Labelling in the plot
text(pretty(xLim)[!pretty(xLim) == 0], -.5, pretty(xLim)[!pretty(xLim) == 0])
text(-.5, pretty(yLim)[!pretty(yLim) == 0], pretty(yLim)[!pretty(yLim) == 0])
text(max(pretty(xLim)), -1, "Theoretical Quantiles", font = 2, adj = 1)
text(-5, min(pretty(yLim)), "Sample Quantiles", font = 2, adj = 0, srt = 90)

qqline(theResiduals, lwd=3, col ="gray75")
```
```{r}
xLim <- range(pretty(range(theModel.Clean$fitted.values)))
yLim <- range(pretty(range(theResiduals)))
plot(x = NULL, y = NULL, xlim = xLim, ylim = yLim, axes = FALSE, ann = FALSE)
axis(1, at = pretty(xLim), lwd = 2, tck = -.015, labels = FALSE, pos=0)
axis(1, at = pretty(xLim), lwd = 2, tck = .015, labels = FALSE, pos=0)
axis(2, at = pretty(yLim), lwd = 2, tck = .015, labels = FALSE)
axis(2, at = pretty(yLim), lwd = 2, tck = -.015, labels = FALSE)
mtext(pretty(yLim),2,at = pretty(yLim), line = .4, las = 1)
mtext("Residuals", 2, font = 2, line= 1.5)

text(pretty(xLim), -.5, pretty(xLim))
text(max(pretty(xLim)), -1, "Fitted Values", font =2, adj=1)
points(theModel$fitted.values, theResiduals, pch = 21, col = colours[2], bg = colours[1], cex = 1.5)
```


```{r}
summary(theModel.Clean)
```

```{r}
theModelS <- summary(theModel.Clean)
```

```{r}
set.seed(42)
theSlopes <- matrix(NA, nrow = 1000, ncol = 6)
for (i in 1:1000){
  IDX <- sample(1:dim(airbnb.Train2)[1], 1000, replace = TRUE)
  sampleDF <- airbnb.Train2[IDX,]
  theModel.temp <- lm( log_price ~ neighbourhood_group + room_type + number_of_reviews + availability_365, data = sampleDF)
  theSlopes[i,] <- theModel.temp$coefficients[2:7]
}

par(lwd = 2)
CI <- apply(theSlopes, 2, quantile, c(0.025, .975))
MM <- apply(theSlopes, 2, mean)
yLimit <- range(pretty(range(CI)))
plot(x = NULL,y = NULL, xlim = c(1,6), ylim = yLimit, axes = FALSE, ann = FALSE)
axis(2, at = pretty(yLimit), lwd = 2, tck = -.015,labels = FALSE)
mtext(pretty(yLimit),2,at =pretty(yLimit), line =.4, las =1)

abline(h=0, col = colours[1], lty = 2, lwd = 2)
arrows(1:6, CI[1,], 1:6, CI[2,], length = .02, angle = 90, code = 3, col = colours[1], lwd = 2)
points(1:6, MM, pch = 19, col = colours[1])
mtext(theVariables, 1, at = 1:6)

```

```{r}
theModel.Final <- lm(log_price ~ minimum_nights + longitude +latitude + number_of_reviews + availability_365,data=airbnb.Train2)
```

```{r}
calculateR2 <- function(yPred, y){
  ssError <- sum((yPred-y)**2)
  ssTotal <- sum((y-mean(y))**2)
  R2 <- 1-ssError/ssTotal
  return(R2)
}
```

```{r}
yPred <- predict(theModel.Final, newdata = airbnb.Test)
calculateR2(yPred, airbnb.Test$log_price)
```

```{r}
calculateRMSerror <- function(yPred, y){
  ssError <- sum((yPred-y)**2)
  rmsE <- sqrt(ssError/length(y))
  return(rmsE)
}
```

```{r}
calculateRMSerror(yPred, airbnb.Test$log_price)
```

```{r}
par(pty = "s")
xLim <- range(pretty(range(yPred, airbnb.Test$log_price)))
yLim <- xLim
plot(x = NULL, y = NULL, xlim = xLim, ylim = yLim, axes = FALSE, ann = FALSE)
axis(1, at = pretty(xLim), lwd = 2, tck = -.015, labels = FALSE)
axis(2, at = pretty(yLim), lwd = 2, tck = -.015, labels = FALSE)
mtext(pretty(xLim),1, at = pretty(xLim), line = .3)
mtext(pretty(xLim),2, at = pretty(xLim), line = .4, las = 1)
mtext("Predicted log10 Price", 2, font = 2, line = 1.5)
mtext("Real log10 Price", 1, font = 2, line = 1.5)
mtext("Long/Lat Model", 3, font = 2, adj = 0)
abline(a = 0, b = 1, lwd = 3, col = "gray75")
# polygon(c(4,4,9,4), c(4,9,9,4), col = rgb(.75,.75,.75,.75,.3), border=NA)

points(airbnb.Test$log_price,yPred, pch = 21, col = colours[2], bg = colours[1], cex = 1.5)

r2 <- calculateR2(yPred, airbnb.Test$log_price)
rmsE <- calculateRMSerror(yPred,airbnb.Test$log_price)

legend("topleft", legend = c(bquote(RMS[error]~ "="~.(round(rmsE,3))), bquote(R^2~"="~.(round(r2,3)))), bty = "n")
```
#model comparision 

```{r}
 theModel1 <- lm(log_price ~ neighbourhood_group + room_type + number_of_reviews + availability_365, data = airbnb.Train2)
summary(theModel)
```
```{r}
calculateAIC <- function(yPred, y,p){
  n <- length(y)
  SSerror <- sum((yPred-y)**2)
  AIC <- 2*p + n*log(SSerror/n)
  return(AIC)
}

adjustedR2 <- function(yPred, y, p){
  r2 <- calculateR2(yPred, y)
  n <- length(y)
  r2adj <- 1-(1-r2)*((n-1)/(n-p-1))
  return(r2adj)
}
```

```{r}
yPred.Original <- predict(theModel.Final, newdata = airbnb.Test)
yPred.neighbourhood <- predict(theModel1, newdata = airbnb.Test)

AIC <- calculateAIC(yPred.Original, airbnb.Test$log_price, length(theModel.Final$coefficients))
AIC[2] <- calculateAIC(yPred.neighbourhood, airbnb.Test$log_price, length(theModel1$coefficients))

R2Adj <- adjustedR2(yPred.Original, airbnb.Test$log_price, length(theModel.Final$coefficients))
R2Adj[2] <- adjustedR2(yPred.neighbourhood, airbnb.Test$log_price, length(theModel1$coefficients))
```

```{r}
data.frame(AIC = AIC, R2Adjusted = R2Adj, row.names = c("orginal Model", "neighborhood  Model"))
```
```{r}
par(pty = "s", mfcol = c(1,2))
xLim <- range(pretty(range(yPred.Original, yPred.neighbourhood, airbnb.Test$log_price)))
yLim <- xLim
plot(x = NULL, y = NULL, xlim = xLim, ylim = yLim, axes = FALSE, ann = FALSE)
axis(1, at = pretty(xLim), lwd = 2, tck = -.015, labels = FALSE)
axis(2, at = pretty(yLim), lwd = 2, tck = -.015, labels = FALSE)
mtext(pretty(xLim),1, at = pretty(xLim), line = .3)
mtext(pretty(xLim),2, at = pretty(xLim), line = .4, las = 1)
mtext("Predicted log10 Price", 2, font = 2, line = 1.5)
mtext("Real log10 Price", 1, font = 2, line = 1.5)
mtext("Long/Lat Model", 3, font = 2, adj = 0)
abline(a = 0, b = 1, lwd = 3, col = "gray75")

points(airbnb.Test$log_price,yPred.Original, pch = 21, col = colours[2], bg = colours[1], cex = 1.5)

legend("topleft", legend = c(paste0("AIC = ", round(AIC[1],2)), bquote(R^2~"="~.(round(R2Adj[1],3)))), bty = "n")


plot(x = NULL, y = NULL, xlim = xLim, ylim = yLim, axes = FALSE, ann = FALSE)
axis(1, at = pretty(xLim), lwd = 2, tck = -.015, labels = FALSE)
axis(2, at = pretty(yLim), lwd = 2, tck = -.015, labels = FALSE)
mtext(pretty(xLim),1, at = pretty(xLim), line = .3)
mtext(pretty(xLim),2, at = pretty(xLim), line = .4, las = 1)
mtext("Predicted log10 Price", 2, font = 2, line = 1.5)
mtext("Real log10 Price", 1, font = 2, line = 1.5)
mtext("neighborhood Model", 3, font = 2, adj = 0)
abline(a = 0, b = 1, lwd = 3, col = "gray75")

points(airbnb.Test$log_price,yPred.neighbourhood, pch = 21, col = colours[2], bg = colours[1], cex = 1.5)

legend("topleft", legend = c(paste0("AIC = ", round(AIC[2],2)), bquote(R^2~"="~.(round(R2Adj[2],3)))), bty = "n")
```
```{r}
yPred.neighbourhood <- predict(theModel1, newdata = airbnb.Validate)
R2Adj.Final <- adjustedR2(yPred.neighbourhood, airbnb.Validate$log_price, length(theModel1$coefficients))
R2Adj.Final
```

